---
title: "BenchmarkGlacTrendScript_v1"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/Desktop/SperryWx/SperryReanalysis/Trend/CalibratedSols')
```


```{r include=FALSE}
######################################Benchmark Glacier Trend Analysis######################################
library(chron)
library(zoo)
library(lubridate)
library(pastecs)
library(ggplot2)
library(gghighlight)
library(ggthemes)
library(tidyr)
library(plotrix)
library(reshape)
library(ggpubr)
library(plyr)
library(dplyr)
library(mice)
library(Kendall)
library(trend)
library(reshape2)
library(lmtest)
library(car)
library(outliers)
library(gvlma)
library(modifiedmk)
library(gitlabr)
library(RCurl)
library(astsa)
library(TSA)
library(gitgadget)
library(usethis)
library(tseries)
library(gdata)
library(grid)
library(gridExtra)
library(extrafont)
library(LaplacesDemon)
library(robustbase)
library(robust)
library(pracma)
library(egg)
library(lemon)
library(patternplot)
loadfonts(device="pdf",quiet = T)
```



```{r include=FALSE}
setwd("~/Desktop/SperryWx/SperryReanalysis/Trend/CalibratedSols")
files = list.files(pattern="*.csv")
allData <- lapply(files, function(.file){
  
  dat<-read.csv(.file, header=T)
  dat$id<-as.character(.file)
  
  dat    # return the dataframe
})
# combine into a single dataframe
D1 <- do.call(rbind.fill, allData)
D1$id <- sub("_Glacier.*", "", D1$id)

#keep pertinent columns
D2 <- D1[,c(1:4,17)]

#remove 2018
D2 <- D2[!D2$Year == "2018",]

Gulk1 <- subset(D2, D2$id == "Gulkana")
Wolv1 <- subset(D2, D2$id == "Wolverine")
SC1 <- subset(D2, D2$id == "South_Cascade")
Sper1 <- subset(D2, D2$id == "Sperry")
LC1 <- subset(D2, D2$id == "Lemon_Creek")

```
#Raw Annual and Seasonal Plots

```{r echo=FALSE, warning=FALSE,fig.width=8.5}
#Gulkana raw plot
cols2=c("Ba"="#999999","Bw"="#E69F00","Bs"="#56B4E9")
gulk.plot1 <- ggplot(Gulk1)+
  geom_line(aes(x=Year,y=Ba_mwe,colour="Ba"),size=1)+
  geom_line(aes(x=Year,y=Bw_mwe,colour="Bw"),size=1)+
  geom_line(aes(x=Year,y=Bs_mwe,colour="Bs"),size=1)+
  geom_smooth(aes(x=Year,y=Bw_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Bs_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Ba_mwe), method="lm")+
  xlab("Year")+
  ylab("m w.e.")+
  scale_x_continuous(breaks = seq(1965,2018,5))+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(legend.position = "right")+
  labs(title="Gulkana")+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="dashed"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=cols2,breaks=c("Ba","Bw","Bs"),labels=c(expression(B[a]),expression(B[w]),expression(B[s])))
gulk.plot1

#Wolverine raw plot
wolv.plot1 <- ggplot(Wolv1)+
  geom_line(aes(x=Year,y=Ba_mwe,colour="Ba"),size=1)+
  geom_line(aes(x=Year,y=Bw_mwe,colour="Bw"),size=1)+
  geom_line(aes(x=Year,y=Bs_mwe,colour="Bs"),size=1)+
  geom_smooth(aes(x=Year,y=Bw_mwe))+
  geom_smooth(aes(x=Year,y=Bs_mwe))+
  geom_smooth(aes(x=Year,y=Ba_mwe))+
  xlab("Year")+
  ylab("m w.e.")+
  scale_x_continuous(breaks = seq(1965,2018,5))+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(legend.position = "right")+
  labs(title="Wolverine")+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="dashed"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=cols2,breaks=c("Ba","Bw","Bs"),labels=c(expression(B[a]),expression(B[w]),expression(B[s])))
wolv.plot1

#SC raw plot
sc.plot1 <- ggplot(SC1)+
  geom_line(aes(x=Year,y=Ba_mwe,colour="Ba"),size=1)+
  geom_line(aes(x=Year,y=Bw_mwe,colour="Bw"),size=1)+
  geom_line(aes(x=Year,y=Bs_mwe,colour="Bs"),size=1)+
  geom_smooth(aes(x=Year,y=Bw_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Bs_mwe),method = "loess")+
  geom_smooth(aes(x=Year,y=Ba_mwe), method="lm")+
  xlab("Year")+
  ylab("m w.e.")+
  scale_x_continuous(breaks = seq(1960,2018,5))+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(legend.position = "right")+
  labs(title="South Cascade")+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="dashed"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=cols2,breaks=c("Ba","Bw","Bs"),labels=c(expression(B[a]),expression(B[w]),expression(B[s])))
sc.plot1

#Sperry raw plot
sper.plot1 <- ggplot(Sper1)+
  geom_line(aes(x=Year,y=Ba_mwe,colour="Ba"),size=1)+
  geom_line(aes(x=Year,y=Bw_mwe,colour="Bw"),size=1)+
  geom_line(aes(x=Year,y=Bs_mwe,colour="Bs"),size=1)+
  geom_smooth(aes(x=Year,y=Bw_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Bs_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Ba_mwe), method="lm")+
  xlab("Year")+
  ylab("m w.e.")+
  scale_x_continuous(breaks = seq(1965,2018,5))+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(legend.position = "right")+
  labs(title="Sperry")+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="dashed"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=cols2,breaks=c("Ba","Bw","Bs"),labels=c(expression(B[a]),expression(B[w]),expression(B[s])))
sper.plot1

#Lemon Creek raw plot
lc.plot1 <- ggplot(LC1)+
  geom_line(aes(x=Year,y=Ba_mwe,colour="Ba"),size=1)+
  geom_line(aes(x=Year,y=Bw_mwe,colour="Bw"),size=1)+
  geom_line(aes(x=Year,y=Bs_mwe,colour="Bs"),size=1)+
  geom_smooth(aes(x=Year,y=Bw_mwe),method = "lm")+
  geom_smooth(aes(x=Year,y=Bs_mwe),method = 'lm')+
  geom_smooth(aes(x=Year,y=Ba_mwe),method = 'loess')+
  xlab("Year")+
  ylab("m w.e.")+
  scale_x_continuous(breaks = seq(1950,2018,5))+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(legend.position = "right")+
  labs(title="Lemon Creek")+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="dashed"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=cols2,breaks=c("Ba","Bw","Bs"),labels=c(expression(B[a]),expression(B[w]),expression(B[s])))
lc.plot1

```
\newpage
#Trend using Mann Kendall
##using Mann-Kendall the following exhibit trends:

##Wolverine annual balance: tau = -0.26, 2-sided pvalue =0.0063099
##Lemon Creek annual balance: tau = -0.426, 2-sided pvalue =4.4747e-07
##Wolverine summer balance: tau = -0.38, 2-sided pvalue =6.2097e-05
##South Cascade summer balance: tau = -0.294, 2-sided pvalue =0.00095174
##Lemon Creek summer balance: tau = -0.381, 2-sided pvalue =0.017053
```{r}
#detrend series
#annual
GGBaDetr <- detrend(Gulk1$Ba_mwe,tt='constant')
WGBaDetr <- detrend(Wolv1$Ba_mwe,tt='constant')
LCBaDetr <- detrend(LC1$Ba_mwe,tt='constant')
SCGBaDetr <- detrend(SC1$Ba_mwe,tt='constant')
SGBaDetr <- detrend(Sper1$Ba_mwe,tt='constant')

#winter
GGBwDetr <- detrend(Gulk1$Bw_mwe,tt='constant')
WGBwDetr <- detrend(Wolv1$Bw_mwe,tt='constant')
LCBwDetr <- detrend(LC1$Bw_mwe,tt='constant')
SCGBwDetr <- detrend(SC1$Bw_mwe,tt='constant')
SGBwDetr <- detrend(Sper1$Bw_mwe,tt='constant')

#summer
GGBsDetr <- detrend(Gulk1$Bs_mwe,tt='constant')
WGBsDetr <- detrend(Wolv1$Bs_mwe,tt='constant')
LCBsDetr <- detrend(LC1$Bs_mwe,tt='constant')
SCGBsDetr <- detrend(SC1$Bs_mwe,tt='constant')
SGBsDetr <- detrend(Sper1$Bs_mwe,tt='constant')

#Mann-Kendall on detrended series
#annual
MannKendall(GGBaDetr)
MannKendall(WGBaDetr)
MannKendall(LCBaDetr)
MannKendall(SCGBaDetr)
MannKendall(SGBaDetr)

#winter
MannKendall(GGBwDetr)
MannKendall(WGBwDetr)
MannKendall(SCGBwDetr)
MannKendall(SGBwDetr)

#summer
MannKendall(GGBsDetr)
MannKendall(WGBsDetr)
MannKendall(SCGBsDetr)
MannKendall(SGBsDetr)


#Mann-Kendall on original series (not detrended)
#annual
MannKendall(Gulk1$Ba_mwe)
MannKendall(Wolv1$Ba_mwe)
MannKendall(SC1$Ba_mwe)
MannKendall(Sper1$Ba_mwe)
MannKendall(LC1$Ba_mwe)

#winter
MannKendall(Gulk1$Bw_mwe)
MannKendall(Wolv1$Bw_mwe)
MannKendall(SC1$Bw_mwe)
MannKendall(Sper1$Bw_mwe)
MannKendall(LC1$Bw_mwe)

#summer
MannKendall(Gulk1$Bs_mwe)
MannKendall(Wolv1$Bs_mwe)
MannKendall(SC1$Bs_mwe)
MannKendall(Sper1$Bs_mwe)
MannKendall(LC1$Bs_mwe)

#amplitude
MannKendall(GG.bamp$amp)
MannKendall(WG.bamp$amp)
MannKendall(SC.bamp$amp)
MannKendall(SG.bamp$amp)

```

```{r include=FALSE}
##################################
#LM for balance
#Gulk
Gulk1.ba.lm <- lmrob(Gulk1$Ba_mwe ~ Gulk1$Year)
Gulk1.bw.lm <- lmrob(Gulk1$Bw_mwe ~ Gulk1$Year)
Gulk1.bs.lm <- lmRob(Gulk1$Bs_mwe ~ Gulk1$Year) #used different package/function so iterations wouldn't reach limit as it does in lmrob

Gulk1.bamp <- Gulk1.bw.lm$residuals + abs(Gulk1.bs.lm$residuals)
Gulk1.bamp.lm <- lmrob(Gulk1.bamp ~ Gulk1$Year)

#Wolv
Wolv1.ba.lm <- lmrob(Wolv1$Ba_mwe ~ Wolv1$Year)
Wolv1.bw.lm <- lmrob(Wolv1$Bw_mwe ~ Wolv1$Year)
Wolv1.bs.lm <- lmrob(Wolv1$Bs_mwe ~ Wolv1$Year)

Wolv1.bamp <- Wolv1.bw.lm$residuals + abs(Wolv1.bs.lm$residuals)
Wolv1.bamp.lm <- lmrob(Wolv1.bamp ~ Wolv1$Year)

#SC
SC1.ba.lm <- lmrob(SC1$Ba_mwe ~ SC1$Year)
SC1.bw.lm <- lmrob(SC1$Bw_mwe ~ SC1$Year)
SC1.bs.lm <- lmrob(SC1$Bs_mwe ~ SC1$Year)

SC1.bamp <- SC1.bw.lm$residuals + abs(SC1.bs.lm$residuals)
SC1.bamp.lm <- lmrob(SC1.bamp ~ SC1$Year)

#Sperry
Sper1.ba.lm <- lmrob(Sper1$Ba_mwe ~ Sper1$Year)
Sper1.bw.lm <- lmrob(Sper1$Bw_mwe ~ Sper1$Year)
Sper1.bs.lm <- lmrob(Sper1$Bs_mwe ~ Sper1$Year)

Sper1.bamp <- Sper1.bw.lm$residuals + abs(Sper1.bs.lm$residuals)
Sper1.bamp.lm <- lmrob(Sper1.bamp ~ Sper1$Year)

#LC
LC1.ba.lm <- lmrob(LC1$Ba_mwe ~ LC1$Year)
LC1.bw.lm <- lmrob(LC1$Bw_mwe ~ LC1$Year)
LC1.bs.lm <- lmrob(LC1$Bs_mwe ~ LC1$Year)

```
\newpage
#Global Test of Model Assumptions (linear)
##Result below show the following do not meet assumptions for linear model:
##Format = Glacier, balance, linear assumption, p-value (level of sig. 0.05)

##Gulkana, winter balance, skewness, 0.03575
##Wolverine, winter balance, non-linear relationship, <0.0001
##Wolverine, winter balance, skewness, <0.0001
##Wolverine, winter balance, kurtosis, <0.0001
##South Cascade, summer balance, heteroscedasticity, 0.02882
##Lemon Creek, annual balance, non-linear relationship, 0.01164
##Lemon Creek, annual balance, heteroscedasticity, 0.02814
##Lemon Creek, winter balance, non-linear relationship, 0.001933
##Lemon Creek, winter balance, skewness, 0.013766
##Lemon Creek, winter balance, kurtosis, 0.002296

```{r, fig.width=8.5,fig.height=10}
# Global test of model assumptions
#Gulk
gvmodel.ba.gulk <- gvlma(Gulk1.ba.lm) 
summary(gvmodel.ba.gulk)
plot.gvlma(gvmodel.ba.gulk)
gvmodel.bw.gulk <- gvlma(Gulk1.bw.lm) 
summary(gvmodel.bw.gulk)
plot.gvlma(gvmodel.bw.gulk)
gvmodel.bs.gulk <- gvlma(Gulk1.bs.lm) 
summary(gvmodel.bs.gulk)
plot.gvlma(gvmodel.bs.gulk)
#Wolv
gvmodel.ba.wolv <- gvlma(Wolv1.ba.lm) 
summary(gvmodel.ba.wolv)
plot.gvlma(gvmodel.ba.wolv)
gvmodel.bw.wolv <- gvlma(Wolv1.bw.lm) 
summary(gvmodel.bw.wolv)
plot.gvlma(gvmodel.bw.wolv)
gvmodel.bs.wolv <- gvlma(Wolv1.bs.lm) 
summary(gvmodel.bs.wolv)
plot.gvlma(gvmodel.bs.wolv)
#SC
gvmodel.ba.sc <- gvlma(SC1.ba.lm) 
summary(gvmodel.ba.sc)
plot.gvlma(gvmodel.ba.sc)
gvmodel.bw.sc <- gvlma(SC1.bw.lm) 
summary(gvmodel.bw.sc)
plot.gvlma(gvmodel.bw.sc)
gvmodel.bs.sc <- gvlma(SC1.bs.lm) 
summary(gvmodel.bs.sc)
plot.gvlma(gvmodel.bs.sc)
#Sperry
gvmodel.ba.sper <- gvlma(Sper1.ba.lm) 
summary(gvmodel.ba.sper)
plot.gvlma(gvmodel.ba.sper)
gvmodel.bw.sper <- gvlma(Sper1.bw.lm) 
summary(gvmodel.bw.sper)
plot.gvlma(gvmodel.bw.sper)
gvmodel.bs.sper <- gvlma(Sper1.bs.lm) 
summary(gvmodel.bs.sper)
plot.gvlma(gvmodel.bs.sper)
#LC
gvmodel.ba.LC <- gvlma(LC1.ba.lm) 
summary(gvmodel.ba.LC)
plot.gvlma(gvmodel.ba.LC)
gvmodel.bw.LC <- gvlma(LC1.bw.lm) 
summary(gvmodel.bw.LC)
plot.gvlma(gvmodel.bw.LC)
gvmodel.bs.LC <- gvlma(LC1.bs.lm) 
summary(gvmodel.bs.LC)
plot.gvlma(gvmodel.bs.LC)

#for those balance time series that exhibit trends, but don't meet linear trend assumptions (only two), let's try....
#LC annual
#fit loess (based on plot above)
LC1.ba.loess <- loess(LC1$Ba_mwe ~ LC1$Year)
#fit 1st order diff model (based on residuals below indicating a 1st order difference correlation)
LC1.diff <- diff(LC1$Ba_mwe)
plot(LC1.diff)
LC1.diff <- as.data.frame(LC1.diff)

#SC summer
#fit loess (based on plot above)
SC1.bs.loess <- loess(SC1$Bs_mwe ~ SC1$Year)

#fit into one df
LC1.df <- cbind(LC1$Year,LC1.ba.lm$residuals,LC1.ba.loess$residuals)
LC1.df <- as.data.frame(LC1.df)
colnames(LC1.df) <- c("Year","lm.resid","loess.resid")
ggplot()+
  geom_line(data=LC1.df,aes(y=lm.resid,x=Year),color="red",size=2)+
  geom_line(data=LC1.df,aes(y=loess.resid,x=Year),color="blue")
```

\newpage
#Testing residuals for normality for each glacier annual and seasonal balances
```{r}
#test for normality of residuals
#Ba
jarque.bera.test(Gulk1.ba.lm$residuals)
jarque.bera.test(Wolv1.ba.lm$residuals)
jarque.bera.test(SC1.ba.lm$residuals)
jarque.bera.test(Sper1.ba.lm$residuals)
jarque.bera.test(LC1.ba.lm$residuals)
jarque.bera.test(LC1.ba.loess$residuals)
hist(LC1.ba.loess$residuals)
hist(LC1.diff)
#bw
jarque.bera.test(Gulk1.bw.lm$residuals)
jarque.bera.test(Wolv1.bw.lm$residuals)
jarque.bera.test(SC1.bw.lm$residuals)
jarque.bera.test(Sper1.bw.lm$residuals)
jarque.bera.test(LC1.bw.lm$residuals)
#bs
jarque.bera.test(Gulk1.bs.lm$residuals)
jarque.bera.test(Wolv1.bs.lm$residuals)
jarque.bera.test(SC1.bs.lm$residuals)
jarque.bera.test(Sper1.bs.lm$residuals)
jarque.bera.test(LC1.bs.lm$residuals)
jarque.bera.test(SC1.bs.loess$residuals)

```

\newpage
#Check for persistence/serial correlation of residuals

```{r, fig.width=8.5,fig.height=10}
#check for serial correlation
#Gulk
acf(Gulk1.ba.lm$residuals)
acf(Gulk1.bw.lm$residuals)
acf(Gulk1.bs.lm$residuals)
lag1.plot(Gulk1.ba.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Gulk1.bw.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Gulk1.bs.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
#Wolv
acf(Wolv1.ba.lm$residuals)
acf(Wolv1.bw.lm$residuals)
acf(Wolv1.bs.lm$residuals)
lag1.plot(Wolv1.ba.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Wolv1.bw.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Wolv1.bs.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
#SC
acf(SC1.ba.lm$residuals)
acf(SC1.bw.lm$residuals)
acf(SC1.bs.lm$residuals)
lag1.plot(SC1.ba.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(SC1.bw.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(SC1.bs.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
#Sper
acf(Sper1.ba.lm$residuals)
acf(Sper1.bw.lm$residuals)
acf(Sper1.bs.lm$residuals)
lag1.plot(Sper1.ba.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Sper1.bw.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(Sper1.bs.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
#LC
acf(LC1.ba.lm$residuals)
acf(LC1.bw.lm$residuals)
acf(LC1.bs.lm$residuals)
lag1.plot(LC1.ba.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(LC1.bw.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(LC1.bs.lm$residuals,max.lag = 10,corr = TRUE,smooth = TRUE)
lag1.plot(LC1.diff,max.lag = 10,corr = TRUE,smooth = TRUE)
#check for serial correlation using Durbin Watson
#Gulk
dwtest(Gulk1.ba.lm)
dwtest(Gulk1.bw.lm)
dwtest(Gulk1.bs.lm)
#Wolv
dwtest(Wolv1.ba.lm)
dwtest(Wolv1.bw.lm)
dwtest(Wolv1.bs.lm)
#SC
dwtest(SC1.ba.lm)
dwtest(SC1.bw.lm)
dwtest(SC1.bs.lm)
#Sper
dwtest(Sper1.ba.lm)
dwtest(Sper1.bw.lm)
dwtest(Sper1.bs.lm)
#LC
dwtest(LC1.ba.lm)
dwtest(LC1.bw.lm)
dwtest(LC1.bs.lm)
dwtest(LC1.ba.loess)
```
###Based on results acf plots and Durbin Watson test, annual balance at Lemon Creek Glacier shows persistence in residuals of linear model of annual balance.

\newpage
#Seasonal correlations
```{r}
cor.test(Gulk1.bw.lm$residuals,Gulk1.bs.lm$residuals)
cor.test(Wolv1.bw.lm$residuals,Wolv1.bs.lm$residuals)
cor.test(Sper1.bw.lm$residuals,Sper1.bs.lm$residuals)
cor.test(SC1.bw.lm$residuals,SC1.bs.lm$residuals)
cor.test(LC1.bw.lm$residuals,LC1.bs.lm$residuals)
```

\newpage
#Interannual Variability
##Standard Deviations
```{r}
#Gulk
Gulk.sda <- sd(Gulk1.ba.lm$residuals,na.rm = T);Gulk.sda
Gulk.sdw <- sd(Gulk1.bw.lm$residuals,na.rm = T);Gulk.sdw
Gulk.sds <- sd(Gulk1.bs.lm$residuals,na.rm = T);Gulk.sds
Gulk.sdamp <- sd(Gulk1.bamp.lm$residuals,na.rm = T);Gulk.sdamp

#Wolv
Wolv.sda <- sd(Wolv1.ba.lm$residuals,na.rm = T);Wolv.sda
Wolv.sdw <- sd(Wolv1.bw.lm$residuals,na.rm = T);Wolv.sdw
Wolv.sds <- sd(Wolv1.bs.lm$residuals,na.rm = T);Wolv.sds
Wolv.sdamp <- sd(Wolv1.bamp.lm$residuals,na.rm = T);Wolv.sdamp

#SC
SC.sda <- sd(SC1.ba.lm$residuals,na.rm = T);SC.sda
SC.sdw <- sd(SC1.bw.lm$residuals,na.rm = T);SC.sdw
SC.sds <- sd(SC1.bs.lm$residuals,na.rm = T);SC.sds
SC.sdamp <- sd(SC1.bamp.lm$residuals,na.rm = T);SC.sdamp

#Sper
Sper.sda <- sd(Sper1.ba.lm$residuals,na.rm = T);Sper.sda
Sper.sdw <- sd(Sper1.bw.lm$residuals,na.rm = T);Sper.sdw
Sper.sds <- sd(Sper1.bs.lm$residuals,na.rm = T);Sper.sds
Sper.sdamp <- sd(Sper1.bamp.lm$residuals,na.rm = T);Sper.sdamp

#LC
LC.sda <- sd(LC1.ba.lm$residuals,na.rm = T);LC.sda
LC.sdw <- 0;LC.sdw
LC.sds <- 0;LC.sds
LC.sda.loess <- sd(LC1.ba.loess$residuals,na.rm = T);LC.sda.loess
```
\newpage
##mean winter balances
```{r}
Gulk.mubw <- mean(Gulk1$Bw_mwe)
Wolv.mubw <- mean(Wolv1$Bw_mwe)
SC.mubw <- mean(SC1$Bw_mwe)
Sper.mubw <- mean(Sper1$Bw_mwe)

Gulk.R <- Gulk.sdw^2/(Gulk.sdw^2 + Gulk.sds^2)
Wolv.R <- Wolv.sdw^2/(Wolv.sdw^2 + Wolv.sds^2)
SC.R <- SC.sdw^2/(SC.sdw^2 + SC.sds^2)
Sper.R <- Sper.sdw^2/(Sper.sdw^2 + Sper.sds^2)
```
\newpage
#Standard Deviation 
```{r,echo=FALSE,fig.width=8.5}
sd.df <- as.data.frame(rbind(Gulk.sda,Wolv.sda,LC.sda,SC.sda,Sper.sda,Gulk.sdw,Wolv.sdw,LC.sdw,SC.sdw,Sper.sdw,Gulk.sds,Wolv.sds,LC.sds,SC.sds,Sper.sds))
sd.df$Name <- c('GG','WG','LCG','SCG','SG','GG','WG','LCG','SCG','SG','GG','WG','LCG','SCG','SG')
sd.df$Balance <- c("Annual","Annual","Annual","Annual","Annual","Winter","Winter","Winter","Winter","Winter", "Summer","Summer","Summer","Summer","Summer")
sd.df$Name <- ordered(sd.df$Name, levels = c("GG","WG","LCG","SCG","SG"))

#colors
group.colors <- c('GG'='#984ea3','WG'='#377eb8','LCG'='#4daf4a','SCG'='#ff7f00','SG'='#e41a1c') #purple, blue, green, orange, red

#plot
sd.bar <- ggplot(sd.df,aes(fill=Name, y=V1,x=Balance))+
  geom_bar(position = "dodge",stat = "identity")+
  xlab(" ")+
  ylab("Standard Deviation")+
  theme_bw()+
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=10))+
  theme(text=element_text(size=18),
        panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(colour="grey", linetype="solid"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=15),
        axis.text.y = element_text(color="black", 
                                   size=15))+
  theme(legend.position = "right")+
  scale_fill_manual(values = group.colors)+
  annotate("text", x=0.64,y=0.03,color="black",label="GG")+
  annotate("text", x=0.82,y=0.03,color="black",label="WG")+
  annotate("text", x=1.0,y=0.03,color="black",label="LCG")+
  annotate("text", x=1.18,y=0.03,color="black",label="SCG")+
  annotate("text", x=1.36,y=0.03,color="black",label="SG")+
  annotate("text", x=1.64,y=0.03,color="black",label="GG")+
  annotate("text", x=1.82,y=0.03,color="black",label="WG")+
  annotate("text", x=2.00,y=0.03,color="black",label="LCG")+
  annotate("text", x=2.18,y=0.03,color="black",label="SCG")+
  annotate("text", x=2.36,y=0.03,color="black",label="SG")+
  annotate("text", x=2.64,y=0.03,color="black",label="GG")+
  annotate("text", x=2.82,y=0.03,color="black",label="WG")+
  annotate("text", x=3,y=0.03,color="black",label="LCG")+
  annotate("text", x=3.18,y=0.03,color="black",label="SCG")+
  annotate("text", x=3.36,y=0.03,color="black",label="SG")
sd.bar
ggsave(sd.bar,file="AllGlacSD_new3.tiff",path='~/Desktop/SperryWx/SperryReanalysis/Trend/TrendPlots',width = 8.5, height = 11, units = "in", device="tiff",dpi = 300)
ggsave(sd.bar,file="AllGlacSD_new3.jpg",path='~/Desktop/SperryWx/SperryReanalysis/Trend/TrendPlots',width = 8.5, height = 11, units = "in", device="jpeg",dpi = 300)
```
\newpage
#Trend
##estimate delta
```{r}
##annual balance
#yearly
dGGa <- first(Gulk1.ba.lm$fitted.values) - last(Gulk1.ba.lm$fitted.values) # m yr^-1
dWGa <- first(Wolv1.ba.lm$fitted.values) - last(Wolv1.ba.lm$fitted.values) # m yr^-1
dSCGa <- first(SC1.ba.lm$fitted.values) - last(SC1.ba.lm$fitted.values) # m yr^-1
dSGa <- first(Sper1.ba.lm$fitted.values) - last(Sper1.ba.lm$fitted.values) # m yr^-1
dLCGa <- first(LC1.ba.lm$fitted.values) - last(LC1.ba.lm$fitted.values) # m yr^-1
dLCGa.loess <- first(LC1.ba.loess$fitted.values) - last(LC1.ba.loess$fitted.values) # m yr^-1

#decadal
ddGGa <- (first(Gulk1.ba.lm$fitted.values) - last(Gulk1.ba.lm$fitted.values))/length(Gulk1$Year)*10 # m yr^-1 decade^-1

ddWGa <- (first(Wolv1.ba.lm$fitted.values) - last(Wolv1.ba.lm$fitted.values))/length(Wolv1$Year)*10 # m yr^-1 decade^-1

ddSCGa <- (first(SC1.ba.lm$fitted.values) - last(SC1.ba.lm$fitted.values))/length(SC1$Year)*10 # m yr^-1 decade^-1

ddSGa <- (first(Sper1.ba.lm$fitted.values) - last(Sper1.ba.lm$fitted.values))/length(Sper1$Year)*10 # m yr^-1 decade^-1

ddLCGa <- (first(LC1.ba.lm$fitted.values) - last(LC1.ba.lm$fitted.values))/length(LC1$Year)*10 # m yr^-1 decade^-1
ddLCGa.loess <- (first(LC1.ba.loess$fitted.values) - last(LC1.ba.loess$fitted.values))/length(LC1$Year)*10 # m yr^-1 decade^-1
#delta.a <- combine(dWGa,dGGa,dLCGa,dSCGa,dSGa);delta.a

##winter balance
#yearly
dGGw <- first(Gulk1.bw.lm$fitted.values) - last(Gulk1.bw.lm$fitted.values) # m yr^-1
dWGw <- first(Wolv1.bw.lm$fitted.values) - last(Wolv1.bw.lm$fitted.values) # m yr^-1
dSCGw <- first(SC1.bw.lm$fitted.values) - last(SC1.bw.lm$fitted.values) # m yr^-1
dSGw <- first(Sper1.bw.lm$fitted.values) - last(Sper1.bw.lm$fitted.values) # m yr^-1
dLCGw <- first(LC1.bw.lm$fitted.values) - last(LC1.bw.lm$fitted.values) # m yr^-1

#decadal
ddGGw <- (first(Gulk1.bw.lm$fitted.values) - last(Gulk1.bw.lm$fitted.values))/length(Gulk1$Year)*10 # m yr^-1 decade^-1

ddWGw <- (first(Wolv1.bw.lm$fitted.values) - last(Wolv1.bw.lm$fitted.values))/length(Wolv1$Year)*10 # m yr^-1 decade^-1

ddSCGw <- (first(SC1.bw.lm$fitted.values) - last(SC1.bw.lm$fitted.values))/length(SC1$Year)*10 # m yr^-1 decade^-1

ddSGw <- (first(Sper1.bw.lm$fitted.values) - last(Sper1.bw.lm$fitted.values))/length(Sper1$Year)*10 # m yr^-1 decade^-1

ddLCGw <- (first(LC1.bw.lm$fitted.values) - last(LC1.bw.lm$fitted.values))/length(LC1$Year)*10 # m yr^-1 decade^-1


##summer balance
#yearly
dGGs <- first(Gulk1.bs.lm$fitted.values) - last(Gulk1.bs.lm$fitted.values) # m yr^-1
dWGs <- first(Wolv1.bs.lm$fitted.values) - last(Wolv1.bs.lm$fitted.values) # m yr^-1
dSCGs <- first(SC1.bs.lm$fitted.values) - last(SC1.bs.lm$fitted.values) # m yr^-1
dSGs <- first(Sper1.bs.lm$fitted.values) - last(Sper1.bs.lm$fitted.values) # m yr^-1
dLCGs <- first(LC1.bs.lm$fitted.values) - last(LC1.bs.lm$fitted.values) # m yr^-1

#decadal
ddGGs <- (first(Gulk1.bs.lm$fitted.values) - last(Gulk1.bs.lm$fitted.values))/length(Gulk1$Year)*10 # m yr^-1 decade^-1

ddWGs <- (first(Wolv1.bs.lm$fitted.values) - last(Wolv1.bs.lm$fitted.values))/length(Wolv1$Year)*10 # m yr^-1 decade^-1

ddSCGs <- (first(SC1.bs.lm$fitted.values) - last(SC1.bs.lm$fitted.values))/length(SC1$Year)*10 # m yr^-1 decade^-1

ddSGs <- (first(Sper1.bs.lm$fitted.values) - last(Sper1.bs.lm$fitted.values))/length(Sper1$Year)*10 # m yr^-1 decade^-1

ddLCGs <- (first(LC1.bs.lm$fitted.values) - last(LC1.bs.lm$fitted.values))/length(LC1$Year)*10 # m yr^-1 decade^-1


##amplitude
dGGamp <- first(Gulk1.bamp.lm$fitted.values) - last(Gulk1.bamp.lm$fitted.values) # m yr^-1
dWGamp <- first(Wolv1.bamp.lm$fitted.values) - last(Wolv1.bamp.lm$fitted.values) # m yr^-1
dSCGamp <- first(SC1.bamp.lm$fitted.values) - last(SC1.bamp.lm$fitted.values) # m yr^-1
dSGamp <- first(Sper1.bamp.lm$fitted.values) - last(Sper1.bamp.lm$fitted.values) # m yr^-1

#signal to noise ratio
#annual
snrs_a <- as.data.frame(c(dGGa/Gulk.sda, dWGa/Wolv.sda, dSCGa/SC.sda, dSGa/Sper.sda, dLCGa/LC.sda));snrs_a
snrs_a$names <- c("Gulk","Wolv","SC","Sper","LC");snrs_a
colnames(snrs_a) <- c("snr","names");snrs_a
#winter
snrs_w <- as.data.frame(c(dGGw/Gulk.sdw, dWGw/Wolv.sdw, dSCGw/SC.sdw, dSGw/Sper.sdw))
snrs_w$names <- c("Gulk","Wolv","SC","Sper")
colnames(snrs_w) <- c("snr","names");snrs_w
#summer
snrs_s <- as.data.frame(c(dGGs/Gulk.sds, dWGs/Wolv.sds, dSCGs/SC.sds, dSGs/Sper.sds))
snrs_s$names <- c("Gulk","Wolv","SC","Sper")
colnames(snrs_s) <- c("snr","names");snrs_s


##t-test (Medwedeff and Roe)
#Roe's t-test
#annual
WGta = dWGa/Wolv.sda*(sqrt((length(Wolv1$Year)-2)/12));WGta
WGpta = 1-pt(WGta,length(Wolv1$Year)-1);WGpta

GGta = dGGa/Gulk.sda*(sqrt((length(Gulk1$Year)-2)/12))
GGpta = 1-pt(GGta,length(Gulk1$Year)-1);GGpta

LCGta = dLCGa/LC.sda*(sqrt((length(LC1$Year)-2)/12))
LCGpta = 1-pt(LCGta,length(LC1$Year)-1);LCGpta

SCGta = dSCGa/SC.sda*(sqrt((length(SC1$Year)-2)/12))
SCGpta = 1-pt(SCGta,length(SC1$Year)-1);SCGpta

SGta = dSGa/Sper.sda*(sqrt((length(Sper1$Year)-2)/12))
SGpta = 1-pt(SGta,length(Sper1$Year)-1);SGpta

#winter
WGtw = dWGw/Wolv.sdw*(sqrt((length(Wolv1$Year)-2)/12))
WGptw = 1-pt(WGtw,length(Wolv1$Year)-1)
WGptw

GGtw = dGGw/Gulk.sdw*(sqrt((length(Gulk1$Year)-2)/12));
GGptw = 1-pt(GGtw,length(Gulk1$Year)-1)
GGptw

SCGtw = dSCGw/SC.sdw*(sqrt((length(SC1$Year)-2)/12));
SCGptw = 1-pt(SCGtw,length(SC1$Year)-1)
SCGptw

SGtw = dSGw/Sper.sdw*(sqrt((length(Sper1$Year)-2)/12));
SGptw = 1-pt(SGtw,length(Sper1$Year)-1)
SGptw

#summer
WGts = dWGs/Wolv.sds*(sqrt((length(Wolv1$Year)-2)/12));
WGpts = 1-pt(WGts,length(Wolv1$Year)-1)
WGpts

GGts = dGGs/Gulk.sds*(sqrt((length(Gulk1$Year)-2)/12));
GGpts = 1-pt(GGts,length(Gulk1$Year)-1)
GGpts

SCGts = dSCGs/SC.sds*(sqrt((length(SC1$Year)-2)/12));
SCGpts = 1-pt(SCGts,length(SC1$Year)-1)
SCGpts

SGts = dSGs/Sper.sds*(sqrt((length(Sper1$Year)-2)/12));
SGpts = 1-pt(SGts,length(Sper1$Year)-1)
SGpts

#amplitude
WGtamp = (dWGamp/Wolv.sdamp)*(sqrt((length(Wolv1$Year)-2)/12));
WGptamp = 1-pt(WGtamp,length(Wolv1$Year)-1)
WGptamp

GGtamp = dGGamp/Gulk.sdamp*(sqrt((length(Gulk1$Year)-2)/12));
GGptamp = 1-pt(GGtamp,length(Gulk1$Year)-1)
GGptamp

SCGtamp = dSCGamp/SC.sdamp*(sqrt((length(SC1$Year)-2)/12));
SCGptamp = 1-pt(SCGtamp,length(SC1$Year)-1)
SCGptamp

SGtamp = dSGamp/Sper.sdamp*(sqrt((length(Sper1$Year)-2)/12));
SGptamp = 1-pt(SGtamp,length(Sper1$Year)-1)
SGptamp

t_matrix <-as.data.frame(rbind(WGpta,WGptw,WGpts,WGptamp,GGpta,GGptw,GGpts,GGptamp,LCGpta,SCGpta,SCGptw,SCGpts,SGpta,SGptw,SGpts,SGptamp))
print(t_matrix)


#SNR
#annual
WG_gamma_a = dWGa/Wolv.sda*(1/length(Wolv1$Year))*10
GG_gamma_a = dGGa/Gulk.sda*(1/length(Gulk1$Year))*10
LCG_gamma_a = dLCGa/LC.sda*(1/length(LC1$Year))*10
SCG_gamma_a = dSCGa/SC.sda*(1/length(SC1$Year))*10
SG_gamma_a = dSGa/Sper.sda*(1/length(Sper1$Year))*10

#winter
WG_gamma_w = dWGw/Wolv.sdw*(1/length(Wolv1$Year))*10
GG_gamma_w = dGGw/Gulk.sdw*(1/length(Gulk1$Year))*10
SCG_gamma_w = dSCGw/SC.sdw*(1/length(SC1$Year))*10
SG_gamma_w = dSGw/Sper.sdw*(1/length(Sper1$Year))*10

#summer
WG_gamma_s = dWGs/Wolv.sds*(1/length(Wolv1$Year))*10
GG_gamma_s = dGGs/Gulk.sds*(1/length(Gulk1$Year))*10
SCG_gamma_s = dSCGs/SC.sds*(1/length(SC1$Year))*10
SG_gamma_s = dSGs/Sper.sds*(1/length(Sper1$Year))*10
```
\newpage
#Time Series of standardized mass balance anomalies 
```{r,echo=FALSE,fig.width=8.5}
#a little renaming of variables
D3 <- D2
D3$id[D3$id == "Gulkana"] <- "GG"
D3$id[D3$id == "Wolverine"] <- "WG"
D3$id[D3$id == "Lemon_Creek"] <- "LCG"
D3$id[D3$id == "South_Cascade"] <- "SCG"
D3$id[D3$id == "Sperry"] <- "SG"
#create df of all amp
#not necessary for plotting
# GG.yrs <- seq(1966,2018,by=1)
# GG.bamp <- as.data.frame(cbind(GG.yrs,Gulk1.bamp))
# GG.bamp$id <- "Gulkana"
# colnames(GG.bamp) <- c("Year","amp","id")
# 
# WG.yrs <- seq(1966,2018,1)
# WG.bamp <- as.data.frame(cbind(WG.yrs,Wolv1.bamp))
# WG.bamp$id <- "Wolverine"
# colnames(WG.bamp) <- c("Year","amp","id")
# 
# SC.yrs <- seq(1959,2018,1)
# SC.bamp <- as.data.frame(cbind(SC.yrs,SC1.bamp))
# SC.bamp$id <- "SCG"
# colnames(SC.bamp) <- c("Year","amp","id")
# 
# SG.yrs <- seq(2005,2017,1)
# SG.bamp <- as.data.frame(cbind(SG.yrs,Sper1.bamp))
# SG.bamp$id <- "SP"
# colnames(SG.bamp) <- c("Year","amp","id")
# 
# glac.bamp <- rbind(GG.bamp,WG.bamp,SC.bamp,SG.bamp)

#plotting with facets
D4 <- D3  
D5 <- melt(D4, id.vars = c("Year","id"))
D5$variable <- ordered(D5$variable, levels = c("Ba_mwe", "Bw_mwe", "Bs_mwe"))
D5$variable <- factor(D5$variable, labels = c("(a)", "(b)","(c)"))
D5$id <- ordered(D5$id, levels = c("GG","WG","LCG","SCG","SG"))
group.colors <- c('GG'='#984ea3','WG'='#377eb8','LCG'='#4daf4a','SCG'='#ff7f00','SG'='#e41a1c') #purple, blue, green, orange, red

#for adding a,b,c to panels
dat_text <- data.frame(
  label = c("a", "b", "c"),
  variable   = c("(a)", "(b)", "(c)"),
  x     = c(1952, 1952, 1952),
  y     = c(3.5,6.5,0.5))

#plot
#####fix y axis a-1 label
#####add panel labels
D5.plot <- ggplot(data = D5,aes(x=Year,y=value,group=id))+
  geom_smooth(aes(color=id),linetype="dashed",method="lm",se=F,size=1.5)+ #aes(linetype=id)...if we need each line to be different
  #scale_linetype_manual(values = c("twodash","longdash","dotdash","dotted","dashed"))+
  geom_line(aes(color=id),stat="smooth",data = subset(D5,variable == "(a)" & id == "GG"),method = "lm",se=F,size=2,alpha=0.75)+
  #geom_line(stat="smooth",data = subset(D5,variable == "(a)" & id == "WG"),method = "lm",se=F,size=2,alpha=0.75)+
  geom_line(aes(color=id),stat="smooth",alpha=0.75,data = subset(D5,variable == "(a)" & id == "LCG"),method = "lm",se=F,size=2)+
  geom_line(aes(color=id),stat="smooth",alpha=0.75,data = subset(D5,variable == "(c)" & id == "SCG"),method = "lm",se=F,size=2)+
  geom_line(aes(color=id),stat="smooth",alpha=0.75,data = subset(D5,variable == "(c)" & id == "WG"),method = "lm",se=F,size=2)+
  geom_line(aes(color=id),stat="smooth",alpha=0.75,data = subset(D5,variable == "(c)" & id == "GG"),method = "lm",se=F,size=2)+
  geom_point(aes(fill=id,shape=id),size=2)+
  scale_shape_manual(values = c(21,22,23,24,25))+
  xlab("Time (Years)")+
  #ylab("Mass Balance (m w.e. a-1)")+
  ylab(expression(Mass~Balance~Anomaly~(m~w.e.~a^-1)))+
  scale_x_continuous(breaks = seq(1950,2018,10))+
  theme_bw()+
  theme(legend.text = element_text(size=12))+
  theme(text=element_text(size=18, family = "Arial"),
        panel.grid.major.y = element_line(colour="grey", linetype="solid"), 
        panel.grid.major.x = element_line(colour="grey", linetype="solid"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top")+
  theme(axis.text.x = element_text(color="black", 
                                   size=12),
        axis.text.y = element_text(color="black", 
                                   size=18))+
  scale_color_manual(values=group.colors)+
  scale_fill_manual(values = group.colors)+
  theme(legend.position = "right",legend.background = element_blank())+
  theme(strip.text.y=element_blank(),strip.background = element_blank())+
  geom_text(data=dat_text,mapping=aes(x=x,y=y,label=label,group=NULL,color=NULL),show.legend =FALSE,size=4,fontface="bold")+
  facet_wrap(~variable, ncol = 1,scales="free_y",strip.position = "right")+
  theme(legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size=5)))
D5.plot
D5.plot <- reposition_legend(D5.plot,position="left",panel = 'panel-1-2')

#save
ggsave(D5.plot,file="BaBwBsAmp_GlacAll.tiff",path='~/Desktop/SperryWx/SperryReanalysis/Trend/TrendPlots',width = 8.5, height = 11, units = "in", device="tiff",dpi = 300)
ggsave(D5.plot,file="BaBwBsAmp_GlacAll.jpg",path='~/Desktop/SperryWx/SperryReanalysis/Trend/TrendPlots',width = 8.5, height = 11, units = "in", device="jpeg",dpi = 300)
```

```{r,eval=F,echo=F,include=F}


```

